Pervasive duplication of tumor suppressor genes preceded parallel evolution

### 2020-06-03 paper_PLOS/paper_PLOS_draft.Rmd
Fixed Figures and converted to bookdown-based knitting, need to fix HTML and WORD outputs

---
title: Pervasive duplication of tumor suppressor genes preceded parallel evolution
  of large bodied Atlantogenatans
author:
- name: Juan Manuel Vazquez
  affiliation: "University of Chicago"
- affiliation: "SUNY Buffalo"
  name: Vincent J Lynch
address:
- address: Department of Human Genetics, 920 East 58th St, Chicago, IL, 60637
  code: "University of Chicago"
- address: Department of Biological Sciences, 551 Cooke Hall, Buffalo NY, 14260
  code: "SUNY Buffalo"
output:
  bookdown::pdf_book:
    base_format: rticles::elsevier_article
    number_sections: yes
    includes:
      in_header: preamble.tex
  pdf_document: default
  word_document: default
  workflowr::wflow_html: default
csl: plos.csl
editor_options:
  chunk_output_type: console
bibliography: rbhb_paenungulates.bib
abstract: "Cancer is an intrinsic disease of multicellular organisms. Within a species,
  the size of an animal, - correlated with the individual's number of cells - and
  its lifespan - correlated with increasing cellular damage over time - positively
  correlate with the risk any individual has to form tumors. Between species, however,
  we do not observe any correlation between size, lifespan, and cancer, a phenomenon
  that referred to as _Peto's Paradox_. Elephants are a particularly interesting member
  of this class of paradoxical animals, since they are a set of large species deeply
  nested in a clade of smaller species, indicating a recent gain of size. Recent work
  has identified several individual cases of gene duplicates contributing to the increased
  cancer resistance of elephants, which suggests that duplication of tumor suppressor
  genes may play a more general role in mediating Peto's Paradox by increasing cancer
  resistance in large, long-lived species. \nBy using a Reciprocal Best-Hit BLAT search
  approach, we investigated copy numbers of all protein-coding genes in _Atlantogenatan_
  genomes to see if there is any correlation between the copy number of duplicates
  and changes body size along the phylogenetic tree. From an initial set of 18,011
  protein-coding genes in hg38, we identified a median of 13,880 genes in _Atlantogenatan_
  genomes, of which a median of 940 genes are duplicated. We find that, just as body
  size fluctuates throughout _Atlantogenata_, genes involved in tumor suppressor pathways
  are also duplicated throughout the phylogenetic tree. Extant species of elephants,
  however, show active transcription of both canonical and duplicated copies of tummor
  suppressors that duplicated prior to and during their sudden increase in body size,
  suggesting that the duplication of tumor suppressor genes facilitates the evolution
  of increased body size by compensating for the increased cancer risk. \n"

@@

```{r Libraries and functions, include = F}
knitr::opts_chunk$set(eval=T, echo=F, message = F, warning = F, out.width="6in",
                      fig.retina = 2, fig.align = "left", dpi = 300)

@@

```{r Plot Gene Copy Trees, include=F}
p.ml.tree.increasing.geneCounts.withSize <- ml.tree.increasing.geneCounts.withSize %>%
  as_tibble %>% mutate(Spc.short = Spc.short %>% str_c('bold("', ., '")')) %>% 
  as.treedata() %>% 
  ggtree(
    .,
    branch.length = "none",
    #continuous=T,
    aes(
      color=lnSize,
      fill=lnSize
    ),
    size=2
    ) +
  scale_color_viridis_c(guide=F) +
  scale_fill_viridis_c(limits=c(size.scale.min, size.scale.max), guide = guide_colorbar(title="Body size (ln(g))", title.position = "left", title.vjust=1,title.theme = element_text(face="bold"))) +
  geom_tiplab(aes(label=Spc.short), offset = 0.4, color="black", size=4, parse=T)+
  geom_tiplab(aes(label=not.me(lnSize, N.Genes, threshold = 7)), geom="label",  offset=-0.5, color="black", size=2, fontface="bold")+ 
  geom_tiplab(aes(label=yes.me(lnSize, N.Genes, threshold = 7)), geom="label",  offset=-0.5, color="grey", size=2)+ 
  geom_nodelab(aes(x=branch, label=not.me(lnSize, N.Genes, threshold = 7)), geom="label", nudge_y = -0.1, nudge_x = -0.5, color="black", size=2, fontface="bold")+ 
  geom_nodelab(aes(x=branch, label=yes.me(lnSize, N.Genes, threshold = 7)), geom="label", nudge_y = -0.1, nudge_x = -0.5, color="grey", size=2, fontface="bold")+ 
  geom_nodelab(aes(x=branch, label=not.me(lnSize, label, threshold = 7)), geom="label", nudge_y = 0.5, nudge_x = -0.5, color="black", size=2, fontface="bold")+ 
  geom_nodelab(aes(x=branch, label=yes.me(lnSize, label, threshold = 7)), geom="label", nudge_y = 0.5, nudge_x = -0.5, color="grey", size=2, fontface="bold")+ 
  geom_nodelab2(aes(subset=(label=="Root"), label = "Atlantogenata", x=branch), geom="label",
               size=2,
               color="grey")+
  theme_tree() + 
  xlim(-0.5,13) +
  xlab("Rate of Change, Body Size") + 
  theme(
    legend.position="bottom",
    # panel.background = element_rect(fill = "lightgrey"),
    text = element_text(face = "plain", 
                        colour = "black", lineheight = 0.9, 
                        hjust = 0.5, vjust = 0.5, angle = 0, margin = margin(), 
                        debug = FALSE), 
    axis.text.x = element_text(size = 4, 
                               colour = "black", face = "bold"),
    axis.title = element_text(size = 6, 
                              colour = "black", face = "bold"),
    plot.title = element_text(size = 6, 
                              colour = "black", lineheight = 1, face = "bold"), 
    legend.title = element_text(size = 6, 
                                face = "bold", colour = "black"), 
    legend.text = element_text(size = 6, 
                               face = "plain", colour = "black")
  )
# p.ml.tree.increasing.geneCounts.withSize

p.ml.tree.increasing.geneCounts.withSize.withBL <- ml.tree.increasing.geneCounts.withSize %>% 
  ggtree(
    .,
    # branch.length = "none",
    #continuous=T,
    aes(
      color=lnSize,
      fill=lnSize
    ),
    size=2
    ) +
  scale_color_viridis_c(guide=F) +
  scale_fill_viridis_c(limits=c(size.scale.min, size.scale.max), guide = guide_colorbar(title="Body size (ln(g))", title.position = "left", title.vjust=1,title.theme = element_text(face="bold"))) +
  geom_tiplab(aes(label=Species.strict), 
              offset = 0.02, size=3,
              color="black")+
  geom_tiplab(aes(label=not.me(lnSize, N.Genes)), geom="label",  
              offset=0, size=3,
              color="black")+ 
  geom_tiplab(aes(label=yes.me(lnSize, N.Genes)), geom="label",  
              offset=0.0, size=3,
              color="grey")+ 
  geom_nodelab(aes(x=branch, label=not.me(lnSize, N.Genes)), geom="label", 
               nudge_y = 0.2, nudge_x = 0, size=3, 
               color="black")+ 
  geom_nodelab(aes(x=branch, label=yes.me(lnSize, N.Genes)), geom="label", 
               nudge_y = -0.2, nudge_x = 0, size=3, 
               color="grey")+ 
  geom_nodelab(aes(x=branch, label=not.me(lnSize, label)), geom="label", 
               nudge_y = 0.2, nudge_x = -0.031, size=3, 
               color="black")+ 
  geom_nodelab(aes(x=branch, label=yes.me(lnSize, label)), geom="label", 
               nudge_y = -0.2, nudge_x = -0.031, size=3, 
               color="grey")+ 
  geom_nodelab2(aes(subset=label=="Root", x=branch), geom="label", 
               size=3, 
               color="grey")+ 
  theme_tree2() + 
  xlim(-0.01,0.45) +
  theme(legend.position = "bottom") +
  xlab("Rate of Change, Body Size")
# p.ml.tree.increasing.geneCounts.withSize.withBL
```

@@

ecnc.onegene.onehit <- 
  ecnc.df1 %>% 
  ggplot(
    aes(
      xmin = x1,
      xmax = x2,
      ymin = y1,
      ymax = y2,
      fill = hit
    )
  ) + 
  geom_rect(color="black") + 
  geom_text(aes(x=(x2+x1)/2, y=(y2+y1)/2, label=hit), color="black")+
  geom_hline(yintercept = 1, color="red", lty="dashed", size=2) + 
  scale_fill_brewer(palette = "Dark2", guide=F) + 
  scale_y_continuous(limits = c(0,10), breaks = 0:10) + 
  scale_x_continuous(limits = c(0,100), breaks = seq(from=0, to=100, by = 25)) + 
  labs(
    x = "Query Length",
    y = "Query Coverage",
    title = "ECNC: One Contiguous Gene, One Hit"
  )

ecnc.twogene.twohit <- 
  ecnc.df2 %>% 
  ggplot(
    aes(
      xmin = x1,
      xmax = x2,
      ymin = y1,
      ymax = y2,
      fill = hit
    )
  ) + 
  geom_rect(color="black") + 
  geom_text(aes(x=(x2+x1)/2, y=(y2+y1)/2, label=hit), color="black")+
  geom_hline(yintercept = 2, color="red", lty="dashed", size=2) + 
  scale_fill_brewer(palette = "Dark2", guide=F) + 
  scale_y_continuous(limits = c(0,10), breaks = 0:10) + 
  scale_x_continuous(limits = c(0,100), breaks = seq(from=0, to=100, by = 25)) + 
  labs(
    x = "Query Length",
    y = "Query Coverage",
    title = "ECNC: Two Contiguous Genes, Two Hits"
  )

ecnc.onegene.threehits <-
  ecnc.df3 %>% 
  ggplot(
    aes(
      xmin = x1,
      xmax = x2,
      ymin = y1,
      ymax = y2,
      fill = hit
    )
  ) + 
  geom_rect(color="black") + 
  geom_text(aes(x=(x2+x1)/2, y=(y2+y1)/2, label=hit), color="black")+
  geom_hline(yintercept = 1, color="red", lty="dashed", size=2) + 
  scale_fill_brewer(palette = "Dark2", guide=F) + 
  scale_y_continuous(limits = c(0,10), breaks = 0:10) + 
  scale_x_continuous(limits = c(0,100), breaks = seq(from=0, to=100, by = 25)) + 
  labs(
    x = "Query Length",
    y = "Query Coverage",
    title = "ECNC: One Fragmented Gene, Multiple Hits"
  )

ecnc.twogene.fourhits <-
  ecnc.df4 %>% 
  ggplot(
    aes(
      xmin = x1,
      xmax = x2,
      ymin = y1,
      ymax = y2,
      fill = hit
    )
  ) + 
  geom_rect(color="black") + 
  geom_text(aes(x=(x2+x1)/2, y=(y2+y1)/2, label=hit), color="black")+
  geom_hline(yintercept = 2, color="red", lty="dashed", size=2) + 
  scale_fill_brewer(palette = "Dark2", guide=F) + 
  scale_y_continuous(limits = c(0,10), breaks = 0:10) + 
  scale_x_continuous(limits = c(0,100), breaks = seq(from=0, to=100, by = 25)) + 
  labs(
    x = "Query Length",
    y = "Query Coverage"#,
    #title = "ECNC: Two Genes, Multiple Hits"
  )

ecnc.onegene.manypartialhits <-
  ecnc.df5 %>% 
  ggplot(
    aes(
      xmin = x1,
      xmax = x2,
      ymin = y1,
      ymax = y2,
      fill = hit
    )
  ) + 
  geom_rect(color="black") + 
  geom_text(aes(x=(x2+x1)/2, y=(y2+y1)/2, label=hit), color="black")+
  geom_hline(yintercept = 1.45, color="red", lty="dashed", size=2) + 
  scale_fill_brewer(palette = "Dark2", guide=F) + 
  scale_y_continuous(limits = c(0,10), breaks = 0:10) + 
  scale_x_continuous(limits = c(0,100), breaks = seq(from=0, to=100, by = 25)) + 
  labs(
    x = "Query Length",
    y = "Query Coverage"
    # title = "ECNC: One Real Gene, Multiple Hits on Conserved Domain"
  )

ecnc.twogene.partialhits <-
  ecnc.df6 %>% 
  ggplot(
    aes(
      xmin = x1,
      xmax = x2,
      ymin = y1,
      ymax = y2,
      fill = hit,
      alpha = transparent,
      linetype = transparent
    )
  ) + 
  geom_rect(color="black") + 
  geom_text(aes(x=(x2+x1)/2, y=(y2+y1)/2, label=label), color="black", inherit.aes = F)+
  geom_hline(yintercept = 2, color="red", lty="dashed", size=2) + 
  scale_fill_brewer(palette = "Dark2", guide=F) + 
  scale_alpha_manual(values = c("0"=1,"1"=0.5), guide=F) + 
  scale_linetype_manual(values = c("0"="solid","1"="dotted"), guide=F) + 
  scale_y_continuous(limits = c(0,10), breaks = 0:10) + 
  scale_x_continuous(limits = c(0,100), breaks = seq(from=0, to=100, by = 25)) + 
  labs(
    x = "Query Length",
    y = "Query Coverage",
    title = "ECNC: Two Real Genes, Unidentified Tails"
  )

@@

  # theme_pubclean()+
  # labs_pubr() + 
  # theme(
  #   text = element_text(face = "plain", 
  #                       colour = "black", lineheight = 0.9, 
  #                       hjust = 0.5, vjust = 0.5, angle = 0, margin = margin(), 
  #                       debug = FALSE), 
  #   axis.text = element_text(size = 6, 
  #                              colour = "black", face = "bold"),
  #   axis.title = element_text(size = 8, 
  #                             colour = "black", face = "bold"),
  #   plot.title = element_text(size = 8, 
  #                             colour = "black", lineheight = 1, face = "bold"), 
  #   legend.text = element_text(size = 6, 
  #                              face = "plain", colour = "black"),
  #   legend.position = "bottom"
  # )
p.cor.GCT.CEGMA <-
  cor.GCT.CEGMA %>%
  ggplot(
    aes(x=Var1, y=Var2, fill = value)
  )+
  geom_tile(color = "white")+
  theme_minimal()+
  coord_fixed()+
  scale_x_discrete(
    position = "top"
  ) +
  scale_y_discrete(position = "right") +
  scale_fill_viridis_c(breaks = c(-1,-0.5, 0, 0.5, 1),labels = c(-1,-0.5, 0, 0.5, 1)) +
  theme_pubclean() +
  theme(
    axis.text.x = element_text(
      angle = -330, vjust = 0,
      size = highlightTextSize(cor.GCT.CEGMA$Var1, cor.toHighlight, sizes = c(6,9)), 
      color = highlightTextColor(cor.GCT.CEGMA$Var1, cor.toHighlight, color = c("black", "red")), 
      hjust = 0
    ),
    axis.text.y = element_text(
      size = highlightTextSize(cor.GCT.CEGMA$Var2, cor.toHighlight, sizes = c(6,9)), 
      color = highlightTextColor(cor.GCT.CEGMA$Var2, cor.toHighlight, color = c("black", "red"))
    ),
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    panel.grid.major = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    axis.ticks = element_blank(),
    legend.justification = c(1, 1),
    legend.position = c(0.4,0.7), #c(0.6, 0.7),
    legend.direction = "horizontal",
    legend.text = element_text(size=9),
    legend.title =  element_text(size=11)
  ) +
  guides(
    fill = guide_colorbar(
      title="Pearson\nCorrelation",
      barwidth = 7, barheight = 1,
      title.position = "top", title.hjust = 0.5,
      draw.llim = T
    )
  )
```
```{r Phylogenetic Correlations}

@@

Our initial hypothesis was that genes which duplicated in lineages that experienced a growth in size would be enriched for membership in tumor suppressor pathways. Thus, we used WebGestalt and its Overrepresentation Analysis (ORA) functionality to determine what pathways were enriched in our duplicated gene sets in each branch relative to our initial query set. For our database, we used Reactome for our primary analysis, but additionally used the KEGG, Panther, Wikipathways, and Wikipathways_cancer databases using WebGestalt ORA. Going through the tree, at no FDR $\leq$ 0.5 is there any significant pathway representation for genes that increased in the branches leading to <!--`r noEnrichmentIncreasing.Reactome[1:length(noEnrichmentIncreasing.Reactome)-1] %>% cat(sep=", ")`, or `r noEnrichmentIncreasing.Reactome[length(noEnrichmentIncreasing.Reactome)]`--> _Afrosoricida_, _Pseudoungulata_, _Elephantina_, _Loxodontini_, _Elephas maximus_, _Mammuthus_, _Loxodonta cyclotis_, _Palaeoloxodon antiquus_, _Loxodona_, _Mammuthus columbi_, _Mammuthus primigenius_, _Procavia capensis_, _Elephantidae_, or _Proboscidea_; furthermore, there are no significant pathway enrichments at FDR<0.5 for genes whose copy number did not change between branches (copy-number-stable) for _Afrotheria__. Note that because _Xenarthra_ was selected as the outgroup, it is not possible to polarize the changes in their gene copy numbers along the tree.

@@

```{r Figure-intro, fig.cap="Atlantogenatans with sequenced genomes, body sizes, and known lifespans [@PuttickAndThomas2015; @HAGR]",  out.width="6in", fig.asp=1.3}
atlantogenata.tree.size.lifespan
```

@@

```{r Figure-CancerSuscept, fig.cap="Cancer succecptibility across Atlantogenata. Branch lengths are set to the magnitude of change in cancer succeptiblity; colors indicate the magnitude and direction of the change.", out.width="6in", fig.asp=1}
p.tree.cancerSucceptibility.Atlantogenata
```

>>>>

---
title: Pervasive duplication of tumor suppressor genes preceded parallel evolution
  of large bodied Atlantogenatans
author:
- name: Juan Manuel Vazquez
  affiliation: "University of Chicago"
- affiliation: "SUNY Buffalo"
  name: Vincent J Lynch
address:
- address: Department of Human Genetics, 920 East 58th St, Chicago, IL, 60637
  code: "University of Chicago"
- address: Department of Biological Sciences, 551 Cooke Hall, Buffalo NY, 14260
  code: "SUNY Buffalo"
always_allow_html: true
output:
  bookdown::html_document2: default
  bookdown::pdf_document2: default
  bookdown::pdf_book:
    base_format: rticles::elsevier_article
    number_sections: yes
    includes:
      in_header: preamble.tex
  bookdown::word_document2: default
csl: plos.csl
editor_options:
  chunk_output_type: console
bibliography: rbhb_paenungulates.bib
abstract: "Cancer is an intrinsic disease of multicellular organisms. Within a species,
  the size of an animal, - correlated with the individual's number of cells - and
  its lifespan - correlated with increasing cellular damage over time - positively
  correlate with the risk any individual has to form tumors. Between species, however,
  we do not observe any correlation between size, lifespan, and cancer, a phenomenon
  that referred to as _Peto's Paradox_. Elephants are a particularly interesting member
  of this class of paradoxical animals, since they are a set of large species deeply
  nested in a clade of smaller species, indicating a recent gain of size. Recent work
  has identified several individual cases of gene duplicates contributing to the increased
  cancer resistance of elephants, which suggests that duplication of tumor suppressor
  genes may play a more general role in mediating Peto's Paradox by increasing cancer
  resistance in large, long-lived species. \nBy using a Reciprocal Best-Hit BLAT search
  approach, we investigated copy numbers of all protein-coding genes in _Atlantogenatan_
  genomes to see if there is any correlation between the copy number of duplicates
  and changes body size along the phylogenetic tree. From an initial set of 18,011
  protein-coding genes in hg38, we identified a median of 13,880 genes in _Atlantogenatan_
  genomes, of which a median of 940 genes are duplicated. We find that, just as body
  size fluctuates throughout _Atlantogenata_, genes involved in tumor suppressor pathways
  are also duplicated throughout the phylogenetic tree. Extant species of elephants,
  however, show active transcription of both canonical and duplicated copies of tummor
  suppressors that duplicated prior to and during their sudden increase in body size,
  suggesting that the duplication of tumor suppressor genes facilitates the evolution
  of increased body size by compensating for the increased cancer risk. \n"

@@

```{r Libraries and functions, include = F}
knitr::opts_chunk$set(eval=T, echo=F, message = F, warning = F, out.width="6in", fig.pos = "H", out.extra = "",
                      fig.retina = 2, dpi = 300)

@@

```{r Plot Gene Copy Trees, include=F}
p.ml.tree.increasing.geneCounts.withSize <- ml.tree.increasing.geneCounts.withSize %>%
  as_tibble %>% mutate(Spc.short = Spc.short %>% str_c('bold("', ., '")')) %>% 
  as.treedata() %>% 
  ggtree(
    .,
    branch.length = "none",
    #continuous=T,
    aes(
      color=lnSize,
      fill=lnSize
    ),
    size=2
    ) +
  scale_color_viridis_c(guide=F) +
  scale_fill_viridis_c(limits=c(size.scale.min, size.scale.max), guide = guide_colorbar(title="Body size (ln(g))", title.position = "left", title.vjust=1,title.theme = element_text(face="bold", size=6))) +
  geom_tiplab(aes(label=Spc.short), offset = 0.4, color="black", size=2, parse=T)+
  geom_tiplab(aes(label=not.me(lnSize, N.Genes, threshold = 7)), geom="label",  offset=-0.5, color="black", size=2, fontface="bold")+ 
  geom_tiplab(aes(label=yes.me(lnSize, N.Genes, threshold = 7)), geom="label",  offset=-0.5, color="grey", size=2)+ 
  geom_nodelab(aes(x=branch, label=not.me(lnSize, N.Genes, threshold = 7)), geom="label", nudge_y = -0.1, nudge_x = -0.5, color="black", size=2, fontface="bold")+ 
  geom_nodelab(aes(x=branch, label=yes.me(lnSize, N.Genes, threshold = 7)), geom="label", nudge_y = -0.1, nudge_x = -0.5, color="grey", size=2, fontface="bold")+ 
  geom_nodelab(aes(x=branch, label=not.me(lnSize, label, threshold = 7)), geom="label", nudge_y = 0.5, nudge_x = -0.5, color="black", size=2, fontface="bold")+ 
  geom_nodelab(aes(x=branch, label=yes.me(lnSize, label, threshold = 7)), geom="label", nudge_y = 0.5, nudge_x = -0.5, color="grey", size=2, fontface="bold")+ 
  geom_nodelab2(aes(subset=(label=="Root"), label = "Atlantogenata", x=branch), geom="label",
               size=2,
               color="grey")+
  theme_tree() + 
  xlim(-0.5,13) +
  xlab("Rate of Change, Body Size") + 
  theme(
    legend.position="bottom",
    # panel.background = element_rect(fill = "lightgrey"),
    text = element_text(face = "plain", 
                        colour = "black", lineheight = 0.9, 
                        hjust = 0.5, vjust = 0.5, angle = 0, margin = margin(), 
                        debug = FALSE), 
    axis.text.x = element_text(size = 4, 
                               colour = "black", face = "bold"),
    axis.title = element_text(size = 6, 
                              colour = "black", face = "bold"),
    plot.title = element_text(size = 6, 
                              colour = "black", lineheight = 1, face = "bold"), 
    legend.title = element_text(size = 6, 
                                face = "bold", colour = "black"), 
    legend.text = element_text(size = 6, 
                               face = "plain", colour = "black")
  )
# p.ml.tree.increasing.geneCounts.withSize

#Prep work
ml.tree.increasing.geneCounts.withSize.table <- ml.tree.increasing.geneCounts.withSize %>% 
  as_tibble %>% 
  mutate(joint.label = str_c(label, ": ", N.Genes))

ml.tree.increasing.geneCounts.withSize.table[which(ml.tree.increasing.geneCounts.withSize.table$label %in% genomeTable.atlantogenata$label),]$joint.label <- ""

ml.tree.increasing.geneCounts.withSize <- ml.tree.increasing.geneCounts.withSize.table %>% as.treedata()

p.ml.tree.increasing.geneCounts.withSize.withBL <- ml.tree.increasing.geneCounts.withSize %>% 
  ggtree(
    .,
    # branch.length = "none",
    #continuous=T,
    aes(
      color=lnSize,
      fill=lnSize
    ),
    size=2
    ) +
  scale_color_viridis_c(guide=F) +
  scale_fill_viridis_c(limits=c(size.scale.min, size.scale.max), guide = guide_colorbar(title="Body size (ln(g))", title.position = "left", title.vjust=1,title.theme = element_text(face="bold"))) +
  geom_tiplab(aes(label=Species.strict), 
              offset = 0.02, size=3,
              color="black")+
  geom_label_repel(aes(
    label=yes.me(lnSize, joint.label, threshold=7)
  ),
    size=3, direction = "x", nudge_x = -0.2, force=10, segment.color = "black",
    color="grey"
  ) + 
  geom_label_repel(aes(
    label=not.me(lnSize, joint.label, threshold=7)
  ),
    size=3, direction = "x",nudge_x = -0.05, force=10, 
    color="black") + 
  geom_tiplab(aes(label=not.me(lnSize, N.Genes, threshold=7)), geom = "label", size=3, color="black") + 
  geom_tiplab(aes(label=yes.me(lnSize, N.Genes, threshold=7)), geom = "label", size=3, color="grey") + 
  geom_nodelab2(aes(subset=(label=="Root"), label = "Atlantogenata", x=branch), geom="label",
               size=2,
               color="black")+
  theme_tree2() + 
  xlim(-0.01,0.4) +
  xlab("Rate of Change, Gene Copy Number") + 
  theme(
    legend.position="bottom",
    # panel.background = element_rect(fill = "lightgrey"),
    text = element_text(face = "plain", 
                        colour = "black", lineheight = 0.9, 
                        hjust = 0.5, vjust = 0.5, angle = 0, margin = margin(), 
                        debug = FALSE), 
    axis.text.x = element_text(size = 4,
                               colour = "black", face = "bold"),
    axis.title = element_text(size = 6, 
                              colour = "black", face = "bold"),
    plot.title = element_text(size = 6, 
                              colour = "black", lineheight = 1, face = "bold"), 
    legend.title = element_text(size = 6, 
                                face = "bold", colour = "black"), 
    legend.text = element_text(size = 6, 
                               face = "plain", colour = "black")
  )

# p.ml.tree.increasing.geneCounts.withSize.withBL
```

@@

ecnc.onegene.onehit <- 
  ecnc.df1 %>% 
  ggplot(
    aes(
      xmin = x1,
      xmax = x2,
      ymin = y1,
      ymax = y2,
      fill = hit
    )
  ) + 
  geom_rect(color="black") + 
  geom_text(aes(x=(x2+x1)/2, y=(y2+y1)/2, label=hit), color="black")+
  geom_hline(yintercept = 1, color="red", lty="dashed", size=2) + 
  scale_fill_brewer(palette = "Dark2", guide=F) + 
  scale_y_continuous(limits = c(0,10), breaks = 0:10) + 
  scale_x_continuous(limits = c(0,100), breaks = seq(from=0, to=100, by = 25)) + 
  labs(
    x = "Query Length",
    y = "Query Coverage"
  )

ecnc.twogene.twohit <- 
  ecnc.df2 %>% 
  ggplot(
    aes(
      xmin = x1,
      xmax = x2,
      ymin = y1,
      ymax = y2,
      fill = hit
    )
  ) + 
  geom_rect(color="black") + 
  geom_text(aes(x=(x2+x1)/2, y=(y2+y1)/2, label=hit), color="black")+
  geom_hline(yintercept = 2, color="red", lty="dashed", size=2) + 
  scale_fill_brewer(palette = "Dark2", guide=F) + 
  scale_y_continuous(limits = c(0,10), breaks = 0:10) + 
  scale_x_continuous(limits = c(0,100), breaks = seq(from=0, to=100, by = 25)) + 
  labs(
    x = "Query Length",
    y = "Query Coverage"
  )

ecnc.onegene.threehits <-
  ecnc.df3 %>% 
  ggplot(
    aes(
      xmin = x1,
      xmax = x2,
      ymin = y1,
      ymax = y2,
      fill = hit
    )
  ) + 
  geom_rect(color="black") + 
  geom_text(aes(x=(x2+x1)/2, y=(y2+y1)/2, label=hit), color="black")+
  geom_hline(yintercept = 1, color="red", lty="dashed", size=2) + 
  scale_fill_brewer(palette = "Dark2", guide=F) + 
  scale_y_continuous(limits = c(0,10), breaks = 0:10) + 
  scale_x_continuous(limits = c(0,100), breaks = seq(from=0, to=100, by = 25)) + 
  labs(
    x = "Query Length",
    y = "Query Coverage"
  )

ecnc.twogene.fourhits <-
  ecnc.df4 %>% 
  ggplot(
    aes(
      xmin = x1,
      xmax = x2,
      ymin = y1,
      ymax = y2,
      fill = hit
    )
  ) + 
  geom_rect(color="black") + 
  geom_text(aes(x=(x2+x1)/2, y=(y2+y1)/2, label=hit), color="black")+
  geom_hline(yintercept = 2, color="red", lty="dashed", size=2) + 
  scale_fill_brewer(palette = "Dark2", guide=F) + 
  scale_y_continuous(limits = c(0,10), breaks = 0:10) + 
  scale_x_continuous(limits = c(0,100), breaks = seq(from=0, to=100, by = 25)) + 
  labs(
    x = "Query Length",
    y = "Query Coverage"
  )

ecnc.onegene.manypartialhits <-
  ecnc.df5 %>% 
  ggplot(
    aes(
      xmin = x1,
      xmax = x2,
      ymin = y1,
      ymax = y2,
      fill = hit
    )
  ) + 
  geom_rect(color="black") + 
  geom_text(aes(x=(x2+x1)/2, y=(y2+y1)/2, label=hit), color="black")+
  geom_hline(yintercept = 1.45, color="red", lty="dashed", size=2) + 
  scale_fill_brewer(palette = "Dark2", guide=F) + 
  scale_y_continuous(limits = c(0,10), breaks = 0:10) + 
  scale_x_continuous(limits = c(0,100), breaks = seq(from=0, to=100, by = 25)) + 
  labs(
    x = "Query Length",
    y = "Query Coverage"
  )

ecnc.twogene.partialhits <-
  ecnc.df6 %>% 
  ggplot(
    aes(
      xmin = x1,
      xmax = x2,
      ymin = y1,
      ymax = y2,
      fill = hit,
      alpha = transparent,
      linetype = transparent
    )
  ) + 
  geom_rect(color="black") + 
  geom_text(aes(x=(x2+x1)/2, y=(y2+y1)/2, label=label), color="black", inherit.aes = F)+
  geom_hline(yintercept = 2, color="red", lty="dashed", size=2) + 
  scale_fill_brewer(palette = "Dark2", guide=F) + 
  scale_alpha_manual(values = c("0"=1,"1"=0.5), guide=F) + 
  scale_linetype_manual(values = c("0"="solid","1"="dotted"), guide=F) + 
  scale_y_continuous(limits = c(0,10), breaks = 0:10) + 
  scale_x_continuous(limits = c(0,100), breaks = seq(from=0, to=100, by = 25)) + 
  labs(
    x = "Query Length",
    y = "Query Coverage"
  )

@@

  # theme_pubclean()+
  # labs_pubr() + 
  # theme(
  #   text = element_text(face = "plain", 
  #                       colour = "black", lineheight = 0.9, 
  #                       hjust = 0.5, vjust = 0.5, angle = 0, margin = margin(), 
  #                       debug = FALSE), 
  #   axis.text = element_text(size = 6, 
  #                              colour = "black", face = "bold"),
  #   axis.title = element_text(size = 8, 
  #                             colour = "black", face = "bold"),
  #   plot.title = element_text(size = 8, 
  #                             colour = "black", lineheight = 1, face = "bold"), 
  #   legend.text = element_text(size = 6, 
  #                              face = "plain", colour = "black"),
  #   legend.position = "bottom"
  # )
p.cor.GCT.CEGMA <-
  cor.GCT.CEGMA %>%
  ggplot(
    aes(x=Var1, y=Var2, fill = value)
  )+
  geom_tile(color = "white")+
  theme_minimal()+
  coord_fixed()+
  scale_x_discrete(
    position = "top"
  ) +
  scale_y_discrete(position = "right") +
  scale_fill_viridis_c(breaks = c(-1,-0.5, 0, 0.5, 1),labels = c(-1,-0.5, 0, 0.5, 1)) +
  theme_pubclean() +
  theme(
    axis.text.x = element_text(
      angle = -270, vjust = 0,
      size = highlightTextSize(cor.GCT.CEGMA$Var1, cor.toHighlight, sizes = c(6,9)), 
      color = highlightTextColor(cor.GCT.CEGMA$Var1, cor.toHighlight, color = c("black", "red")), 
      hjust = 0
    ),
    axis.text.y = element_text(
      size = highlightTextSize(cor.GCT.CEGMA$Var2, cor.toHighlight, sizes = c(6,9)), 
      color = highlightTextColor(cor.GCT.CEGMA$Var2, cor.toHighlight, color = c("black", "red"))
    ),
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    panel.grid.major = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    axis.ticks = element_blank(),
    legend.justification = c(1, 1),
    legend.position = c(0.4,0.7), #c(0.6, 0.7),
    legend.direction = "horizontal",
    legend.text = element_text(size=9),
    legend.title =  element_text(size=11)
  ) +
  guides(
    fill = guide_colorbar(
      title="Pearson\nCorrelation",
      barwidth = 7, barheight = 1,
      title.position = "top", title.hjust = 0.5,
      draw.llim = T
    )
  )
```
```{r Phylogenetic Correlations}

@@

Our initial hypothesis was that genes which duplicated in lineages that experienced a growth in size would be enriched for membership in tumor suppressor pathways. Thus, we used WebGestalt and its Overrepresentation Analysis (ORA) functionality to determine what pathways were enriched in our duplicated gene sets in each branch relative to our initial query set. For our database, we used Reactome for our primary analysis, but additionally used the KEGG, Panther, Wikipathways, and Wikipathways_cancer databases using WebGestalt ORA. Going through the tree, at no FDR $\leq$ 0.5 is there any significant pathway representation for genes that increased in the branches leading to <!--`r noEnrichmentIncreasing.Reactome[1:length(noEnrichmentIncreasing.Reactome)-1] %>% cat(sep=", ")`, or `r noEnrichmentIncreasing.Reactome[length(noEnrichmentIncreasing.Reactome)]`--> _Afrosoricida_, _Pseudoungulata_, _Elephantina_, _Loxodontini_, _Elephas maximus_, _Mammuthus_, _Loxodonta cyclotis_, _Palaeoloxodon antiquus_, _Loxodona_, _Mammuthus columbi_, _Mammuthus primigenius_, _Procavia capensis_, _Elephantidae_, or _Proboscidea_; furthermore, there are no significant pathway enrichments at FDR<0.5 for genes whose copy number did not change between branches (copy-number-stable) for _Afrotheria_. Note that because _Xenarthra_ was selected as the outgroup, it is not possible to polarize the changes in their gene copy numbers along the tree.

@@

(ref:Figure-intro) _Atlantogenatans_ with sequenced genomes, body sizes, and known lifespans [@PuttickAndThomas2015; @HAGR]

```{r Figure-intro, fig.cap="(ref:Figure-intro)",  out.width="6in", fig.asp=1.2}
atlantogenata.tree.size.lifespan
```

@@

(ref:Figure-CancerSuscept) Cancer succecptibility across _Atlantogenata._ Branch lengths are set to the magnitude of change in cancer succeptiblity; colors indicate the magnitude and direction of the change.

```{r Figure-CancerSuscept, fig.cap="(ref:Figure-CancerSuscept)", out.width="6in", fig.asp=1}
p.tree.cancerSucceptibility.Atlantogenata
```

(ref:Figure-GeneDup-Cladogram) Gene duplications occur readily throughout _Atlantogenata_. Shown here is a tree of _Atlantogenatan_ species with genomes, with the number of genes that underwent an increase in copy number overlayed at each node.

```{r Figure-GeneDup-Cladogram, fig.cap="(ref:Figure-GeneDup-Cladogram)", out.height="6in", out.width="6in"}
p.ml.tree.increasing.geneCounts.withSize
```
